---
title: "Statistical Inference Project Part 1"
output:
  pdf_document:
    fig_height: 4
---

First I simulated 1000 averages of 40 expeonetials with lambda = 0.2. 
```{r}
lambda <- 0.2
sim <- replicate(1000,(mean(rexp(40,rate=lambda))))
```

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution. 

We know that theoretical mean is 1/lambda. I calculated the mean of my simulated data and it is very close to the theoretical mean.
```{r}
theoretical.mean <- 1/lambda
theoretical.mean
sim.mean <- mean(sim)
sim.mean
```

2. Show how variable it is and compare it to the theoretical variance of the distribution.

We know the standard deviation for a distrubation of means is S/sqrt(n). So the standard deviation is (1/lambda)/sqrt(40). The variance is just the standard deviation squared. I calculated the variance of my simulated data and it is very close to the theoretical variance.
```{r}
theoretical.sd <- (1/lambda)/sqrt(40)
theoretical.var <- theoretical.sd^2
theoretical.var
sim.var <- var(sim)
sim.var

```

3. Show that the distribution is approximately normal.

I first created a histogram of my simulated data. It appeared symetrical, centered around the mean of 5, and roughly normal. To show this more conclusively, I created a qqplot which compares a distribution (in this case my simulated data) to the normal distribution. A perfectly normal distribution should lie on the line. This shows that my data is very close to a normal distribution. 

```{r, echo=FALSE}
hist(sim)
```


```{r, echo=FALSE}
qqnorm(sim)
qqline(sim)
```

4. Evaluate the coverage of the confidence interval for 1/lambda. (This only needs to be done for the specific value of lambda).

I first created the 95% confidence interval. I then checked to see how many of my means of 40 exponentials fell within this interval. I found that approximately 95% of the means fell within this interval. 
```{r}
conf.interval <- theoretical.mean + c(-1,1)*1.96*theoretical.sd
mean(conf.interval[1] < sim & sim < conf.interval[2])
```